---
title: "Optimisation"
author: " "
date: "`r Sys.Date()`" 
output:
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 5,
                      fig.align = 'center', message = FALSE, warning = FALSE,
                      fig.pos='H', echo=FALSE)

```


```{r lib, echo=FALSE}
# setwd("./script")
library(generain)
library(reshape2)
library(ggplot2)
source("load_libraries.R")
library(kableExtra)
library(extRemes)
library(bbmle)
library(ismev)
library(extRemes)
library(evd)
library(dplyr)
library(latex2exp)

btf_theme <- theme_minimal() +
  theme(axis.text.x = element_text(size =  6, angle = 0),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        title = element_text(size = 10),
        axis.line = element_blank(),  # Remove axis lines
        panel.border = element_blank(),  # Remove plot border
        panel.background = element_rect(fill = "transparent", color = NA),
        # Remove plot background
        axis.text = element_blank(),  # Remove axis text
        axis.ticks = element_blank(),  # Remove axis ticks
        plot.background = element_rect(fill = "transparent", color = NA),
        panel.grid = element_line(color = "#5c595943"))

# my green color
btfgreen <- "#69b3a2"
```

# The $r$-Pareto process

Our rainfall process is $X=\{X_{s, t}, (s, t) \in S \times T\}$, where $S$ is
the spatial domain and $T$ the temporal domain.

We have
$X_{s, t} \mid r(X_{s, t}) > u \xrightarrow{d} Z_{s, t}$ with a risk function 
$r(X_{s, t}) = X_{s_0, t_0}$, a threshold $u>1$ and $Z=\{Z_{s, t}, (s, t) \in S \times T\}$ 
a $r$-Pareto process.

With $W=\{W_{s, t}, (s, t) \in S \times T\}$ the Gaussian process of the 
Brown-Resnick process and its variogram $\gamma$, 
we can define the $r$-Pareto process as
\[
Z_{s, t} = R_{s, t} e^{W_{s, t} - W_{s_0, t_0} - \gamma(s - s_0, t - t_0)}
\]
where $R_{s, t}$ is a random variable following a simple Pareto distribution. 

The variogram $\gamma$ is defined as
\[
\gamma(ds, dt) = \beta_1 |ds|^{\alpha_1} + \beta_2 dt^{\alpha_2}
\]
with $ds = s - s'$, $dt=t-t'$, $\beta_1, \beta_2 > 0$ and $\alpha_1, \alpha_2 \in (0, 1)$.

If we add an advection vector $V = (v_x, v_y)$, the variogram becomes
\[
\gamma(ds, dt) = \beta_1 ||ds - V|dt|||^{\alpha_1} + \beta_2 dt^{\alpha_2}.
\]


# The $r$-Pareto process without advection

## Simulation 

We simulate the $r$-Pareto process with the parameters $\beta_1 = 0.4$,
$\beta_2 = 0.2$, $\alpha_1 = 1.5$, $\alpha_2 = 1$ and without advection. 
We simulate the process on a $5 \times 5$ grid with $30$ time steps and $m=100$ realizations.
We use a conditonal point $s_0 = (1, 1)$ at time $t_0 = 1$.

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Time series for 4 sites of the first realization"}
adv <- c(0, 0) # advection
params <- c(0.4, 0.2, 1.5, 1) # ok verif sur simu
true_param <- c(params, adv)
beta1 <- params[1]
beta2 <- params[2]
alpha1 <- params[3]
alpha2 <- params[4]
ngrid <- 5
spa <- 1:ngrid
nsites <- ngrid^2 # if the grid is squared
temp <- 1:30

# Conditional point
s0 <- c(1, 1)
t0 <- 1

# Number of realizations
nres <- 100

# Simulate the process
# simu <- sim_rpareto(beta1, beta2, alpha1, alpha2, spa, spa, temp, adv, s0,
#                     t0, nres)

if (any(adv < 1)) {
  adv_int <- adv * 10
} else {
  adv_int <- adv
}
adv_str <- sprintf("%02d_%02d", adv_int[1], adv_int[2])

# Save the data
foldername <- paste0("../data/simulations_rpar/sim_", ngrid^2, "s_",
                                length(temp), "t_", adv_str, "/")


if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

# save_simulations(simu, ngrid, nres, folder = foldername,
#         file = paste0("rpar_", ngrid^2, "s_", length(temp), "t"))

# nres <- length(list.files(foldername))
list_simu <- list()
for (i in 1:nres) {
  file_name <- paste0(foldername, "rpar_", ngrid^2, "s_",
                        length(temp), "t_", i, ".csv")
  list_simu[[i]] <- read.csv(file_name)
}

# Plot the first realization
simu_df <- list_simu[[1]]
sites_coords <- generate_grid_coords(ngrid)
par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
plot(simu_df$S1, main = "S1=S0", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S2, main = "S2", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S10, main = "S10", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S20, main = "S20", xlab = "Time", ylab = "Rainfall")
```

```{r together2, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Time series for 4 sites of the all replicates together"}
simu_all <- do.call(rbind, list_simu)

par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
plot(simu_all$S1, main = "S1=S0", xlab = "Time", ylab = "Rainfall")
plot(simu_all$S2, main = "S2", xlab = "Time", ylab = "Rainfall")
plot(simu_all$S10, main = "S10", xlab = "Time", ylab = "Rainfall")
plot(simu_all$S20, main = "S20", xlab = "Time", ylab = "Rainfall")
```

## Estimation of the variogram parameters

```{r, fig.width = 5, fig.height = 5, echo=FALSE}
# df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
#                                         tau_max = 10)
# u <- 1
# list_excesses <- list()
# # par(mfrow=c(5, 2))
# for (i in 1:nres) {
#   list_excesses[[i]] <- empirical_excesses_rpar(list_simu[[i]], u, df_lags,
#                                            threshold = TRUE, t0=t0)
# }

# result <- optim(par = true_param, fn = neg_ll_composite,
#                   list_simu = list_simu,
#                   quantile = u,
#                   df_lags = df_lags,
#                   list_excesses = list_excesses,
#                   hmax = sqrt(17),
#                   s0 = s0,
#                   t0 = t0,
#                   threshold = TRUE,
#                   method = "L-BFGS-B",
#                   lower = c(1e-6, 1e-6, 1e-6, 1e-6, -Inf, -Inf),
#                   upper = c(Inf, Inf, 1.999, 1.999, Inf, Inf),
#                   control = list(parscale = c(1, 1, 1, 1, 1, 1),
#                                  maxit = 10000,
#                                  ndeps = c(1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3)))


# df_result <- data.frame(beta1 =  result$par[1],
#                         beta2 = result$par[2],
#                         alpha1 = result$par[3],
#                         alpha2 = result$par[4],
#                         adv1 = result$par[5],
#                         adv2 = result$par[6])

# df_rmse <- data.frame(beta1 = sqrt((result$par[1] - true_param[1])^2),
#                 beta2 = sqrt((result$par[2] - true_param[2])^2),
#                 alpha1 = sqrt((result$par[3] - true_param[3])^2),
#                 alpha2 = sqrt((result$par[4] - true_param[4])^2),
#                 adv1 = sqrt((result$par[5] - true_param[5])^2),
#                 adv2 = sqrt((result$par[6] - true_param[6])^2))

# df_result <- rbind(df_result, df_rmse)
# rownames(df_result) <- c("estim", "rmse")
# kable(df_result, "latex", booktabs = TRUE,
#         caption = "RMSE for all simulations together")  %>%
#     kable_styling(latex_options = "H",
#         bootstrap_options = c("striped", "hover", "condensed", "responsive"))


# save the result
foldername <- "../data/optim/rpar_noadv/"
if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

name_file <- paste0("result_optim_", nres, "simu_", ngrid^2, 
                                "s_", length(temp), "t_" , adv_str, ".csv")

# write.csv(df_result, paste0(foldername, name_file))
# get csv
df_result <- read.csv(paste0(foldername, name_file))
colnames(df_result) <- c("", "beta1", "beta2", "alpha1", "alpha2", "adv1", "adv2")
kable(df_result, "latex", booktabs = TRUE,
        caption = "Result and RMSE for all replicates together with advection=(0,0)")  %>%
    kable_styling(latex_options = "H",
        bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram of the estimated parameters"}
# Plot the variogram
df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
                                      tau_max = 10)
tau_values <- c(0, 1, 5, 10)
result <- df_result[1, ]

generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Boxplot of the estimated parameters"}
# Plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```

### For multiple sets of replicates

We can look at $M=100$ simulations with $m=100$ replicates each.

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Time series for 4 sites of the first realization"}
adv <- c(0, 0) # advection
params <- c(0.4, 0.2, 1.5, 1) # ok verif sur simu
true_param <- c(params, adv)
beta1 <- params[1]
beta2 <- params[2]
alpha1 <- params[3]
alpha2 <- params[4]
ngrid <- 5
spa <- 1:ngrid
nsites <- ngrid^2 # if the grid is squared
temp <- 1:30

# Conditional point
s0 <- c(1, 1)
t0 <- 1

# Number of realizations
M <- 50
m <- 100
nres <- M * m

# Simulate the process
# simu <- sim_rpareto(beta1, beta2, alpha1, alpha2, spa, spa, temp, adv, s0,
#                     t0, nres)

if (any(adv < 1)) {
  adv_int <- adv * 10
} else {
  adv_int <- adv
}
adv_str <- sprintf("%02d_%02d", adv_int[1], adv_int[2])

# Save the data
foldername <- paste0("../data/simulations_rpar/sim_", ngrid^2, "s_",
                                length(temp), "t_", adv_str, "/")


if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

# save_simulations(simu, ngrid, nres, folder = foldername,
#         file = paste0("rpar_", ngrid^2, "s_", length(temp), "t"))

# nres <- length(list.files(foldername))
list_simu <- list()
for (i in 1:nres) {
  file_name <- paste0(foldername, "rpar_", ngrid^2, "s_",
                        length(temp), "t_", i, ".csv")
  list_simu[[i]] <- read.csv(file_name)
}

# Concatenate all the simulations by m replicates
list_simu_all <- list()
for(i in 1:M) {
  list_simu_all[[i]] <- do.call(rbind, list_simu[((i - 1) * m + 1):(i * m)])
}

# Plot the first simulations combining m replicates
simu_all <- list_simu_all[[1]]
sites_coords <- generate_grid_coords(ngrid)
par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
plot(simu_all$S1, main = "S1=S0", xlab = "Time", ylab = "Rainfall")
plot(simu_all$S2, main = "S2", xlab = "Time", ylab = "Rainfall")
plot(simu_all$S10, main = "S10", xlab = "Time", ylab = "Rainfall")
plot(simu_all$S20, main = "S20", xlab = "Time", ylab = "Rainfall")
```

Optimisation of the variogram parameters for the $M=100$ simulations with $m=100$ replicates each.

```{r optimMsimu, fig.width = 5, fig.height = 5, echo=FALSE}
# df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
#                                         tau_max = 10)

# df_result_all <- data.frame(beta1 = numeric(M),
#                         beta2 = numeric(M),
#                         alpha1 = numeric(M),
#                         alpha2 = numeric(M),
#                         adv1 = numeric(M),
#                         adv2 = numeric(M))
# u <- 1

# for (i in 1:M){
#   # get the m corresponding simulations from list_simu inside a list
#   mreplicates <- list_simu[((i - 1) * m + 1):(i * m)]

#   list_excesses <- list()
#   for (j in 1:m) {
#     replicate <- mreplicates[[j]]
#     list_excesses[[j]] <- empirical_excesses_rpar(replicate, u, df_lags,
#                                             threshold = TRUE, t0=t0)
#   }

#   result <- optim(par = true_param, fn = neg_ll_composite,
#                   list_simu = mreplicates,
#                   quantile = u,
#                   df_lags = df_lags,
#                   list_excesses = list_excesses,
#                   hmax = sqrt(17),
#                   s0 = s0,
#                   t0 = t0,
#                   threshold = TRUE,
#                   method = "L-BFGS-B",
#                   lower = c(1e-6, 1e-6, 1e-6, 1e-6, -Inf, -Inf),
#                   upper = c(Inf, Inf, 1.999, 1.999, Inf, Inf),
#                   control = list(parscale = c(1, 1, 1, 1, 1, 1),
#                                  maxit = 10000,
#                                  ndeps = c(1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3)))

#   df_result_all[i, ] <- result$par

# }

# # get only 50 first rows
# df_result_all <- df_result_all[1:50, ]

# # Boxplot of the results with ggplot for each column
# df_bplot <- df_result_all
# df_bplot <- stack(df_bplot)


# title_str <- paste0("Boxplot of the estimated parameters for the ", M,
#                    " simulations of ", m, " replicates each")

# ggplot(df_bplot, aes(x = ind, y = values)) +
#   geom_boxplot() +
#   labs(title = title_str,
#     x = "Parameters", y = "Estimated values") +
#   theme_minimal() +
#   geom_point(aes(y = true_param[as.numeric(ind)]), color = "red", pch=4) 


# save plot
# foldername <- "../images/optim/rpar_noadv/"
# if (!dir.exists(foldername)) {
#   dir.create(foldername, recursive = TRUE)
# }

# name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
#                                 "s_", length(temp), "t_", adv_str, ".png")

# ggsave(paste0(foldername, name_file), width = 5, height = 5)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Boxplot of the estimated parameters for the 100 simulations of 100 replicates each"}
# get png from folder
foldername <- "../images/optim/rpar_noadv/"
if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
                                "s_", length(temp), "t_", adv_str, ".png")

knitr::include_graphics(paste0(foldername, name_file))
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE}
# df_valid <- get_criterion(df_result_all, true_param)

# save the result
foldername <- "../data/optim/rpar_adv/"
if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

name_file <- paste0("result_optim_valid_", M, "simu_", m, "rep_", ngrid^2,
                                "s_", length(temp), "t_" , adv_str, ".csv")

# write.csv(df_valid, paste0(foldername, name_file))
# get csv
df_result <- read.csv(paste0(foldername, name_file))
colnames(df_result) <- c("", "Mean", "RMSE", "MAE")
kable(df_result, "latex", booktabs = TRUE,
        caption = "Result and RMSE for all replicates together with advection=(0,0)")  %>%
    kable_styling(latex_options = "H",
        bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the mean estimated parameters for the 50 simulations of 100 replicates each"}
# plot variogram
result <- data.frame(par = df_result$Mean)
generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the mean estimated parameters for the 50 simulations of 100 replicates each"}
# plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```

# The $r$-Pareto process with advection

## Simulation

We simulate the $r$-Pareto process with the parameters $\beta_1 = 0.4$,
$\beta_2 = 0.2$, $\alpha_1 = 1.5$, $\alpha_2 = 1$ and the advection vector
$V = (0.5, 0.3)$. We simulate the process on a $5 \times 5$ grid with
$30$ time steps and $100$ realizations. We use a conditonal point $s_0 = (1, 1)$ at
time $t_0 = 1$.

```{r rparetosim1, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Time series for 4 sites of the first realization in the advection direction"}
adv <- c(0.5, 0.3) # advection
params <- c(0.4, 0.2, 1.5, 1) # ok verif sur simu
true_param <- c(params, adv)
beta1 <- params[1]
beta2 <- params[2]
alpha1 <- params[3]
alpha2 <- params[4]
ngrid <- 5
spa <- 1:ngrid
nsites <- ngrid^2 # if the grid is squared
temp <- 1:30

# Conditional point
s0 <- c(1, 1)
t0 <- 1

# Number of realizations
M <- 50
m <- 1000
nres <- M * m

# Simulate the process
simu <- sim_rpareto(beta1, beta2, alpha1, alpha2, spa, spa, temp, adv, s0,
                    t0, nres)

if (any(adv < 1)) {
  adv_int <- adv * 10
} else {
  adv_int <- adv
}
adv_str <- sprintf("%02d_%02d", adv_int[1], adv_int[2])

# Save the data
foldername <- paste0("../data/simulations_rpar/sim_", ngrid^2, "s_",
                                length(temp), "t_", adv_str, "/")


if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

save_simulations(simu, ngrid, nres, folder = foldername,
        file = paste0("rpar_", ngrid^2, "s_", length(temp), "t"))

# nres <- length(list.files(foldername))
list_simu <- list() # first simulation of m replicates
for (i in 1:m) {
  file_name <- paste0(foldername, "rpar_", ngrid^2, "s_",
                        length(temp), "t_", i, ".csv")
  list_simu[[i]] <- read.csv(file_name)
}

list_simuM <- list()
for (i in 1:nres) {
  file_name <- paste0(foldername, "rpar_", ngrid^2, "s_",
                        length(temp), "t_", i, ".csv")
  list_simuM[[i]] <- read.csv(file_name)
}

# Plot the first realization
simu_df <- list_simu[[2]]
sites_coords <- generate_grid_coords(ngrid)
par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
plot(simu_df$S1, main = "S1=S0=(1,1)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S7, main = "S7=(2,2)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S10, main = "S13=(3,3)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S20, main = "S25=(5,5)", xlab = "Time", ylab = "Rainfall")
```

## Optimisation

We want to estimate the parameters $\beta_1$, $\beta_2$, $\alpha_1$ and $\alpha_2$
of the $r$-Pareto process. We use the maximum likelihood estimation with the
composite likelihood method. 

We compute the number of joint excesses for each replicate $i$,  $k^{(i)}_{s, t} = \sum_{t=1}^T \mathbb{1}_{\{X_{s, t} > u, X_{s_0, t_0} > u\}}$
and $k^{(i)}_{s-s_0, t-t_0} \sim Bin(T - t - t_0, \chi(s - s_0, t - t_0))$ with $T$ the number of observations
within a replicate (same for all replicates).


```{r optimrparadv, fig.width = 5, fig.height = 5, echo=FALSE}
df_lags <- get_conditional_lag_vectors(sites_coords, s0, t0, tau_max = 10)

u <- 1
list_excesses <- list()
# par(mfrow=c(5, 2))
for (i in 1:m) {
  list_excesses[[i]] <- empirical_excesses(list_simu[[i]], u, df_lags,
                                     threshold = TRUE, type="rpareto", t0=1)
}

result <- optim(par = true_param, fn = neg_ll_composite,
                  list_simu = list_simu,
                  quantile = u,
                  df_lags = df_lags,
                  list_excesses = list_excesses,
                  hmax = sqrt(17),
                  s0 = s0,
                  t0 = t0,
                  threshold = TRUE,
                  method = "L-BFGS-B",
                  lower = c(1e-6, 1e-6, 1e-6, 1e-6, -Inf, -Inf),
                  upper = c(Inf, Inf, 1.999, 1.999, Inf, Inf),
                  control = list(parscale = c(1, 1, 1, 1, 1, 1),
                                 maxit = 10000,
                                 ndeps = c(1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3)))


df_result <- data.frame(beta1 =  result$par[1],
                        beta2 = result$par[2],
                        alpha1 = result$par[3],
                        alpha2 = result$par[4],
                        adv1 = result$par[5],
                        adv2 = result$par[6])

df_rmse <- data.frame(beta1 = sqrt((result$par[1] - true_param[1])^2),
                beta2 = sqrt((result$par[2] - true_param[2])^2),
                alpha1 = sqrt((result$par[3] - true_param[3])^2),
                alpha2 = sqrt((result$par[4] - true_param[4])^2),
                adv1 = sqrt((result$par[5] - true_param[5])^2),
                adv2 = sqrt((result$par[6] - true_param[6])^2))

df_result <- rbind(df_result, df_rmse)
rownames(df_result) <- c("estim", "rmse")

# save the result
foldername <- "../data/optim/rpar_adv/"
if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

name_file <- paste0("result_optim_", m, "simu_", ngrid^2,
                                "s_", length(temp), "t_", adv_str, ".csv")

# write.csv(df_result, paste0(foldername, name_file))

# get csv
df_result <- read.csv(paste0(foldername, name_file))
kable(df_result, "latex", booktabs = TRUE,
        caption = "Result and RMSE for all replicates together with advection=(0.5,0.3)")  %>%
    kable_styling(latex_options = "H",
        bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

Plot the estimated variogram vs the true theorical variogram.

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the estimated parameters"}
# Plot the variogram
df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
                                      tau_max = 10)
tau_values <- c(0, 1, 5, 10)
result <- data.frame(df_result[1, c("beta1", "beta2", "alpha1", "alpha2",
                                      "adv1", "adv2")])
generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the estimated parameters"}
# Plot the chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```

For $M = 50$ simulations with $m = 100$ replicates each, we have the following results:

```{r optimMsimuadv, fig.width = 5, fig.height = 5, echo=FALSE}
df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
                                      tau_max = 10)

# df_result_all <- data.frame(beta1 = numeric(M),
#                         beta2 = numeric(M),
#                         alpha1 = numeric(M),
#                         alpha2 = numeric(M),
#                         adv1 = numeric(M),
#                         adv2 = numeric(M))
# u <- 1

# for (i in 1:M){
#   # get the m corresponding simulations from list_simu inside a list
#   mreplicates <- list_simuM[((i - 1) * m + 1):(i * m)]

#   list_excesses <- list()
#   for (j in 1:m) {
#     replicate <- mreplicates[[j]]
#     list_excesses[[j]] <- empirical_excesses_rpar(replicate, u, df_lags,
#                                             threshold = TRUE, t0=t0)
#   }

#   result <- optim(par = true_param, fn = neg_ll_composite,
#                   list_simu = mreplicates,
#                   quantile = u,
#                   df_lags = df_lags,
#                   list_excesses = list_excesses,
#                   hmax = sqrt(17),
#                   s0 = s0,
#                   t0 = t0,
#                   threshold = TRUE,
#                   method = "L-BFGS-B",
#                   lower = c(1e-6, 1e-6, 1e-6, 1e-6, -Inf, -Inf),
#                   upper = c(Inf, Inf, 1.999, 1.999, Inf, Inf),
#                   control = list(parscale = c(1, 1, 1, 1, 1, 1),
#                                  maxit = 10000,
#                                  ndeps = c(1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3)))

#   df_result_all[i, ] <- result$par

# }

# save the result
foldername <- "../data/optim/rpar_adv/"
if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

name_file <- paste0("result_optim_all_", M, "simu_", m, "rep_", ngrid^2,
                                "s_", length(temp), "t_", adv_str, ".csv")

# write.csv(df_result_all, paste0(foldername, name_file))

# get csv
df_result_all <- read.csv(paste0(foldername, name_file))
df_result_all <- df_result_all[c("beta1", "beta2", "alpha1", "alpha2", "adv1", "adv2")]
# kable(df_result_all, "latex", booktabs = TRUE,
#         caption = "Result for all simulations together with advection=(0.5,0.3)")  %>%
#     kable_styling(latex_options = "H",
#         bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# df_valid <- get_criterion(df_result_all, true_param)
# name_file <- paste0("valid_optim_", M, "simu_", m, "rep_", ngrid^2,
#                                 "s_", length(temp), "t_" , adv_str, ".csv")

# write.csv(df_valid, paste0(foldername, name_file))

# get csv
df_valid <- read.csv(paste0(foldername, name_file))
kable(df_valid, "latex", booktabs = TRUE,
        caption = "Criterion for all simulations together with advection=(0.5,0.3)")  %>%
    kable_styling(latex_options = "H",
        bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# Boxplot of the results with ggplot for each column
# df_bplot <- df_result_all
# df_bplot <- stack(df_bplot)


# ggplot(df_bplot, aes(x = ind, y = values)) +
#   geom_boxplot() +
#   labs(title = "",
#     x = "Parameters", y = "Estimated values") +
#   theme_minimal() +
#   geom_point(aes(y = true_param[as.numeric(ind)]), color = "red", pch=4) 


# save plot
# foldername <- "../images/optim/rpar_adv/"
# if (!dir.exists(foldername)) {
#   dir.create(foldername, recursive = TRUE)
# }
# # setwd("./script")
# name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
#                                 "s_", length(temp), "t_", adv_str, ".png")

# ggsave(paste0(foldername, name_file), width = 5, height = 5)
```


```{r, fig.width = 7, fig.height = 3, echo=FALSE, fig.cap="Boxplot of the estimated parameters for the 50 simulations of 100 replicates each"}
# save plot
foldername <- "../images/optim/rpar_adv/"

name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
                                "s_", length(temp), "t_", adv_str, ".png")

knitr::include_graphics(paste0(foldername, name_file))
```

For one simulation:

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the estimated parameters for the 30th simulation of 100 replicates each"}
# plot variogram
df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
                                      tau_max = 10)
tau_values <- c(0, 1, 5, 10)
result <- df_result_all[30, ]
generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the estimated parameters for the 30th simulation of 100 replicates each"}
# plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the estimated parameters for the 50th simulation of 100 replicates each"}
# plot variogram
tau_values <- c(0, 1, 5, 10)
result <- df_result_all[50, ]
generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the estimated parameters for the 50th simulation of 100 replicates each"}
# plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the estimated parameters for the third simulation of 100 replicates each"}
# plot variogram
tau_values <- c(0, 1, 5, 10)
result <- df_result_all[3, ]
generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the estimated parameters for the third simulation of 100 replicates each"}
# plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```

With mean estimation for the $M=50$ simulations with $m=100$ replicates each.
```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the mean estimated parameters for the 50 simulations of 100 replicates each"}
# plot variogram
# tau_values <- c(0, 1, 5, 10)
# df_valid <- df_valid[c("mean", "RMSE", "MAE")]
# result <- data.frame(par = df_valid$mean)
# generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the mean estimated parameters for the 50 simulations of 100 replicates each"}
# plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```


## Plot by fixing advection parameters

```{r, fig.width = 5, fig.height = 5, echo=FALSE, eval=FALSE}
df_result_all <- data.frame(beta1 = numeric(M),
                        beta2 = numeric(M),
                        alpha1 = numeric(M),
                        alpha2 = numeric(M))
u <- 1

for (i in 1:M){
  # get the m corresponding simulations from list_simu inside a list
  mreplicates <- list_simuM[((i - 1) * m + 1):(i * m)]

  list_excesses <- list()
  for (j in 1:m) {
    replicate <- mreplicates[[j]]
    list_excesses[[j]] <- empirical_excesses_rpar(replicate, u, df_lags,
                                            threshold = TRUE, t0=t0)
  }

  result <- mle2(neg_ll_composite_par,
              start = list(beta1 = true_param[1],
                           beta2 = true_param[2],
                           alpha1 = true_param[3],
                           alpha2 = true_param[4],
                           adv1 = true_param[5],
                           adv2 = true_param[6]),
              data = list(list_simu = list_simu,
                          quantile = u,
                          df_lags = df_lags,
                          list_excesses = list_excesses,
                          hmax = sqrt(17),
                          s0 = s0,
                          t0 = t0,
                          threshold = TRUE,
                          method = "CG"),
              control = list(maxit = 10000),
              fixed = list(adv1=true_param[5], adv2=true_param[6]))

  df_result_all[i, ] <- result@coef

}


# Boxplot of the results with ggplot for each column
# df_bplot <- df_result_all
# df_bplot <- stack(df_bplot)


# ggplot(df_bplot, aes(x = ind, y = values)) +
#   geom_boxplot() +
#   labs(title = "",
#     x = "Parameters", y = "Estimated values") +
#   theme_minimal() +
#   geom_point(aes(y = true_param[as.numeric(ind)]), color = "red", pch=4) 


# # save plot
# foldername <- "../images/optim/rpar_adv/"
# if (!dir.exists(foldername)) {
#   dir.create(foldername, recursive = TRUE)
# }
# # setwd("./script")
# name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
#                                 "s_", length(temp), "t_fixed_", adv_str, ".png")

# ggsave(paste0(foldername, name_file), width = 5, height = 5)

```

```{r, fig.width = 7, fig.height = 3, echo=FALSE, fig.cap="Boxplot of the estimated parameters for the 50 simulations of 100 replicates each with advection parameters fixed"}
# # save plot
# foldername <- "../images/optim/rpar_adv/"

# name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
#                                 "s_", length(temp), "t_fixed_", adv_str, ".png")

# knitr::include_graphics(paste0(foldername, name_file))
```


## With another advection with $V_x=0$

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Time series for 4 sites of the first realization in the advection direction"}
adv <- c(0, 0.5) # advection
params <- c(0.4, 0.2, 1.5, 1) # ok verif sur simu
true_param <- c(params, adv)
beta1 <- params[1]
beta2 <- params[2]
alpha1 <- params[3]
alpha2 <- params[4]
ngrid <- 5
spa <- 1:ngrid
nsites <- ngrid^2 # if the grid is squared
temp <- 1:30

# Conditional point
s0 <- c(1, 1)
t0 <- 1

# Number of realizations
M <- 50
m <- 100
nres <- M * m

# Simulate the process
# simu <- sim_rpareto(beta1, beta2, alpha1, alpha2, spa, spa, temp, adv, s0,
#                     t0, nres)

if (any(adv < 1)) {
  adv_int <- adv * 10
} else {
  adv_int <- adv
}
adv_str <- sprintf("%02d_%02d", adv_int[1], adv_int[2])

# Save the data
foldername <- paste0("../data/simulations_rpar/sim_", ngrid^2, "s_",
                                length(temp), "t_", adv_str, "/")


if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

# save_simulations(simu, ngrid, nres, folder = foldername,
#         file = paste0("rpar_", ngrid^2, "s_", length(temp), "t"))

# nres <- length(list.files(foldername))
list_simu <- list() # first simulation of m replicates
for (i in 1:m) {
  file_name <- paste0(foldername, "rpar_", ngrid^2, "s_",
                        length(temp), "t_", i, ".csv")
  list_simu[[i]] <- read.csv(file_name)
}

list_simuM <- list()
for (i in 1:nres) {
  file_name <- paste0(foldername, "rpar_", ngrid^2, "s_",
                        length(temp), "t_", i, ".csv")
  list_simuM[[i]] <- read.csv(file_name)
}

# Plot the first realization
simu_df <- list_simu[[2]]
sites_coords <- generate_grid_coords(ngrid)
par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
plot(simu_df$S1, main = "S1=S0=(1,1)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S7, main = "S7=(2,2)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S10, main = "S13=(3,3)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S20, main = "S25=(5,5)", xlab = "Time", ylab = "Rainfall")
```


```{r, fig.width = 5, fig.height = 5, echo=FALSE}
df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
                                      tau_max = 10)

# df_result_all <- data.frame(beta1 = numeric(M),
#                         beta2 = numeric(M),
#                         alpha1 = numeric(M),
#                         alpha2 = numeric(M),
#                         adv1 = numeric(M),
#                         adv2 = numeric(M))
# u <- 1

# for (i in 1:M){
#   # get the m corresponding simulations from list_simu inside a list
#   mreplicates <- list_simuM[((i - 1) * m + 1):(i * m)]

#   list_excesses <- list()
#   for (j in 1:m) {
#     replicate <- mreplicates[[j]]
#     list_excesses[[j]] <- empirical_excesses_rpar(replicate, u, df_lags,
#                                             threshold = TRUE, t0=t0)
#   }

#   result <- optim(par = true_param, fn = neg_ll_composite,
#                   list_simu = mreplicates,
#                   quantile = u,
#                   df_lags = df_lags,
#                   list_excesses = list_excesses,
#                   hmax = sqrt(17),
#                   s0 = s0,
#                   t0 = t0,
#                   threshold = TRUE,
#                   method = "L-BFGS-B",
#                   lower = c(1e-6, 1e-6, 1e-6, 1e-6, -Inf, -Inf),
#                   upper = c(Inf, Inf, 1.999, 1.999, Inf, Inf),
#                   control = list(parscale = c(1, 1, 1, 1, 1, 1),
#                                  maxit = 10000,
#                                  ndeps = c(1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3)))

#   df_result_all[i, ] <- result$par

# }

# save result
foldername <- "../data/optim/rpar_adv/"

name_file <- paste0("result_optim_all_", M, "simu_", m, "rep_", ngrid^2,
                                "s_", length(temp), "t_", adv_str, ".csv")

# write.csv(df_result_all, paste0(foldername, name_file))

# get csv
df_result_all <- read.csv(paste0(foldername, name_file))
df_result_all <- df_result_all[c("beta1", "beta2", "alpha1", "alpha2", "adv1", "adv2")]
# kable(df_result_all, "latex", booktabs = TRUE,
#         caption = "Result for all simulations together with advection=(0,0.5)")  %>%
#     kable_styling(latex_options = "H",
#         bootstrap_options = c("striped", "hover", "condensed", "responsive"))

df_valid <- get_criterion(df_result_all, true_param)
kable(df_valid, "latex", booktabs = TRUE,
        caption = "Criterion for all simulations together with advection=(0,0.5)")  %>%
    kable_styling(latex_options = "H",
        bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

```{r, fig.width = 7, fig.height = 3, echo=FALSE, fig.cap="Boxplot of the estimated parameters for the 50 simulations of 100 replicates each"}
# save plot
# foldername <- "../images/optim/rpar_adv/"

# name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
#                                 "s_", length(temp), "t_", adv_str, ".png")

# Boxplot of the results with ggplot for each column
df_bplot <- df_result_all
df_bplot <- stack(df_bplot)


ggplot(df_bplot, aes(x = ind, y = values)) +
  geom_boxplot() +
  labs(title = "",
    x = "Parameters", y = "Estimated values") +
  theme_minimal() +
  geom_point(aes(y = true_param[as.numeric(ind)]), color = "red", pch=4) 

# # save plot
# foldername <- "../images/optim/rpar_adv/"
# if (!dir.exists(foldername)) {
#   dir.create(foldername, recursive = TRUE)
# }
# # setwd("./script")
# name_file <- paste0("boxplot_optim_", M, "simu_", m, "rep_", ngrid^2,
#                                 "s_", length(temp), "t_", adv_str, ".png")

# ggsave(paste0(foldername, name_file), width = 5, height = 5)

# knitr::include_graphics(paste0(foldername, name_file))

```

For one simulation:

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the estimated parameters for the 30th simulation of 100 replicates each"}
# plot variogram
df_lags <- get_conditional_lag_vectors(sites_coords, true_param, temp, s0, t0,
                                      tau_max = 10)
tau_values <- c(0, 1, 5, 10)
result <- df_result_all[30, ]
generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the estimated parameters for the 30th simulation of 100 replicates each"}
# plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```


```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Variogram for the mean estimated parameters for the 50 simulations of 100 replicates each"}
# plot variogram
tau_values <- c(0, 1, 5, 10)
result <- data.frame(par = df_valid$mean)
generate_variogram_plots(result, df_lags, true_param, tau_values)
```

```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Extremogram for the mean estimated parameters for the 50 simulations of 100 replicates each"}
# plot chi
generate_variogram_plots(result, df_lags, true_param, tau_values, chi = TRUE)
```