---
title: "Simulations"
author: " "
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 5,
                      fig.align = 'center', message = FALSE, warning = FALSE,
                      fig.pos='H', echo=FALSE)

```


```{r lib, echo=FALSE}
# setwd("./script")
library(generain)
library(latex2exp)
library(reshape2)
library(ggplot2)
source("load_libraries.R")
library(kableExtra)
library(extRemes)
library(bbmle)
library(ismev)
library(extRemes)
library(evd)
library(dplyr)

btf_theme <- theme_minimal() +
  theme(axis.text.x = element_text(size =  6, angle = 0),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        title = element_text(size = 10),
        axis.line = element_blank(),  # Remove axis lines
        panel.border = element_blank(),  # Remove plot border
        panel.background = element_rect(fill = "transparent", color = NA),
        # Remove plot background
        axis.text = element_blank(),  # Remove axis text
        axis.ticks = element_blank(),  # Remove axis ticks
        plot.background = element_rect(fill = "transparent", color = NA),
        panel.grid = element_line(color = "#5c595943"))

# my green color
btfgreen <- "#69b3a2"
```

# Simulation Brown-Resnick

Comparison with and without advection on a 2x2 grid, 30 time steps and 2 sites. The 
advection is set to 0.5 in the x direction and 0.2 in the y direction. Seed is fixed.

```{r, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Rainfall simulation with and without advection"}
adv <- c(0.5, 0.2)
true_param <- c(0.4, 0.2, 1.5, 1, adv)
ngrid <- 2
spa <- 1:ngrid
nsites <- ngrid^2 # if the grid is squared
temp <- 1:30
n.BR <- 1

# generate the simulations
set.seed(123)
BR_adv <- sim_BR(true_param[1], true_param[2], true_param[3],
                    true_param[4], spa, spa, temp, adv)
# plot(BR_adv[1, 1, , ], main = "Rainfall simulation with advection")

set.seed(123)
BR_noadv <- sim_BR(true_param[1], true_param[2], true_param[3],
                   true_param[4], spa, spa, temp, adv=c(0, 0))
# plot(BR_noadv[1, 1, , ], main = "Rainfall simulation without advection")

# plot on the same graph with ggplot
df_adv <- data.frame(time = temp,
                     value = BR_adv[1, 1, , ],
                     type = "advection")
df_noadv <- data.frame(time = temp,
                      value = BR_noadv[1, 1, , ],
                      type = "no_advection")

df_plot <- rbind(df_adv, df_noadv)

ggplot(df_plot, aes(x = time, y = value, color = type)) +
  geom_point() +
  labs(title = "",
       x = "Time", y = "Rainfall")
```


## Simulation with 25 sites and 300 time steps without advection

```{r}
adv <- c(0, 0)
true_param <- c(0.4, 0.2, 1.5, 1, adv)
ngrid <- 5
spa <- 1:ngrid
nsites <- ngrid^2 # if the grid is squared
temp <- 1:300

# BR <- sim_BR(true_param[1], true_param[2], true_param[3],
#                     true_param[4], spa, spa, temp, adv)

foldername <- paste0("../data/simulations_BR/sim_", ngrid^2, "s_",
                                length(temp), "t/")
if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

# save_simulations(BR, ngrid, n.BR, folder = foldername,
#         file = paste0("br_", ngrid^2, "s_", length(temp), "t"), forcedind = 1)

# load the simulations
file_path <- paste0(foldername, "br_", ngrid^2, "s_", length(temp),
                "t_", 1, ".csv")
simu_df <- read.csv(file_path)
```


```{r}
# simulation gif
# create_simu_gif(simu_df, true_param, type = "br", forcedtemp = 50)
```


```{r,echo=FALSE, fig.height=5, fig.width=5, fig.cap="Rainfall simulation with 25 sites and 300 time steps"}
# plot the simulation for one site
par(mfrow = c(2, 2), cex = 0.5)
plot(simu_df[, 1], main = "Site 1")
plot(simu_df[, 2], main = "Site 2")
plot(simu_df[, 3], main = "Site 3")
plot(simu_df[, 4], main = "Site 4")
```


```{r, fig.height=4, fig.width=7, fig.cap="Rainfall simulation with 25 sites and 300 time steps on the same plot"}
# plot the simulation at four different sites on 100 time steps
df_plot <- data.frame(time = 1:300,
                      S3 = simu_df[, 3],
                      S4 = simu_df[, 4],
                      S5 = simu_df[, 5])

ggplot(df_plot, aes(x = time)) +
    geom_point(aes(y = S3, color = "S4")) +
    geom_point(aes(y = S4, color = "S5")) +
    geom_point(aes(y = S5, color = "S6")) +
    labs(title = "For four different sites",
         x = "Time", y = "Rainfall") +
    theme_minimal()
```


## Verification of marginal distributions (Gumbel)

```{r, fig.height=5, fig.width=5, fig.cap="QQ-plot of the Gumbel distribution for two sites"}
# qq-plots margins
par(mfrow = c(2, 2))

BR_loc <- simu_df$S4
plot(BR_loc, main = "Site 4")
BR_loc_log <- log(BR_loc)
gumbel.fit <- gum.fit(BR_loc_log, show=FALSE)

mu <- gumbel.fit$mle[1]
sigma <- gumbel.fit$mle[2]

theorical_qgum <- qgumbel(ppoints(BR_loc_log), mu, sigma)

qqplot(BR_loc_log, theorical_qgum, main = "Gumbel Q-Q plot",
       xlab = "Empirical quantiles",
       ylab = "Theoretical quantiles")


BR_loc <- simu_df$S5
plot(BR_loc, main = "Site 5")
BR_loc_log <- log(BR_loc)
gumbel.fit <- gum.fit(BR_loc_log, show=FALSE)

mu <- gumbel.fit$mle[1]
sigma <- gumbel.fit$mle[2]

theorical_qgum <- qgumbel(ppoints(BR_loc_log), mu, sigma)

qqplot(BR_loc_log, theorical_qgum, main = "Gumbel Q-Q plot",
       xlab = "Empirical quantiles",
       ylab = "Theoretical quantiles")
```


## Simulation of r-pareto with advection

The optimisation of the variogram parameters without advection works. So I need to 
focus on the advection part and see if my r-pareto advected simulation works. I will
simulate a 5x5 grid with 30 time steps and 20 realizations.

```{r rparetosim1, echo=FALSE}
adv <- c(0.5, 0.3) # advection
params <- c(0.4, 0.2, 1.5, 1) # ok verif sur simu
true_param <- c(params, adv)
beta1 <- params[1]
beta2 <- params[2]
alpha1 <- params[3]
alpha2 <- params[4]
ngrid <- 5
spa <- 1:ngrid
nsites <- ngrid^2 # if the grid is squared
temp <- 1:30

# Conditional point
s0 <- c(1, 1)
t0 <- 1

# Number of realizations
nres <- 20

# Simulate the process
# simu <- sim_rpareto(beta1, beta2, alpha1, alpha2, spa, spa, temp, adv, s0,
#                     t0, nres)

if (any(adv < 1)) {
  adv_int <- adv * 10
} else {
  adv_int <- adv
}
adv_str <- sprintf("%02d_%02d", adv_int[1], adv_int[2])

# Save the data
foldername <- paste0("../data/simulations_rpar/sim_", ngrid^2, "s_",
                                length(temp), "t_", adv_str, "/")


if (!dir.exists(foldername)) {
  dir.create(foldername, recursive = TRUE)
}

# save_simulations(simu, ngrid, nres, folder = foldername,
#         file = paste0("rpar_", ngrid^2, "s_", length(temp), "t"))

# nres <- length(list.files(foldername))
```


```{r, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Time series for 4 sites of the first realization in the advection direction"}
list_simu <- list()
for (i in 1:nres) {
  file_name <- paste0(foldername, "rpar_", ngrid^2, "s_",
                        length(temp), "t_", i, ".csv")
  list_simu[[i]] <- read.csv(file_name)
}

# Plot the first realization
simu_df <- list_simu[[2]]
sites_coords <- generate_grid_coords(ngrid)
par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
plot(simu_df$S1, main = "S1=S0=(1,1)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S7, main = "S7=(2,2)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S10, main = "S13=(3,3)", xlab = "Time", ylab = "Rainfall")
plot(simu_df$S20, main = "S25=(5,5)", xlab = "Time", ylab = "Rainfall")
```

## Verification of marginal distributions (GPD)

```{r, fig.height=5, fig.width=5, fig.cap="QQ-plot of the GPD distribution for four sites, with quantile at 0.8 and threshold is 0.147"}
# qq-plots margins for 1 site
par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
rpar <- simu_df$S1
threshold <- quantile(rpar, probs = 0.8)
rpar_exc <- rpar[rpar > threshold]
fit_gpd <- gpd.fit(rpar, threshold, show=FALSE)
sigma <- fit_gpd$mle[1]
xi <- fit_gpd$mle[2]

theorical_qgpd <- qgpd(ppoints(rpar_exc), loc=min(rpar_exc), 
                       shape=xi, scale=sigma)

qqplot(rpar_exc, theorical_qgpd, main = "Site 1",
  xlab = "Empirical quantiles",
  ylab = "Theoretical quantiles")

rpar <- simu_df$S3
threshold <- quantile(rpar, probs = 0.8)
rpar_exc <- rpar[rpar > threshold]
fit_gpd <- gpd.fit(rpar, threshold, show=FALSE)
sigma <- fit_gpd$mle[1]
xi <- fit_gpd$mle[2]

theorical_qgpd <- qgpd(ppoints(rpar_exc), loc=min(rpar_exc), 
                       shape=xi, scale=sigma)

qqplot(rpar_exc, theorical_qgpd, main = "Site 3",
  xlab = "Empirical quantiles",
  ylab = "Theoretical quantiles")

rpar <- simu_df$S8
threshold <- quantile(rpar, probs = 0.8)
rpar_exc <- rpar[rpar > threshold]
fit_gpd <- gpd.fit(rpar, threshold)
sigma <- fit_gpd$mle[1]
xi <- fit_gpd$mle[2]

theorical_qgpd <- qgpd(ppoints(rpar_exc), loc=min(rpar_exc), 
                       shape=xi, scale=sigma)

qqplot(rpar_exc, theorical_qgpd, main = "Site 8",
  xlab = "Empirical quantiles",
  ylab = "Theoretical quantiles")

rpar <- simu_df$S10
threshold <- quantile(rpar, probs = 0.8)
rpar_exc <- rpar[rpar > threshold]
fit_gpd <- gpd.fit(rpar, threshold, show=FALSE)
sigma <- fit_gpd$mle[1]
xi <- fit_gpd$mle[2]

theorical_qgpd <- qgpd(ppoints(rpar_exc), loc=min(rpar_exc), 
                       shape=xi, scale=sigma)

qqplot(rpar_exc, theorical_qgpd, main = "Site 10",
  xlab = "Empirical quantiles",
  ylab = "Theoretical quantiles")
```

With the threshold $u=1$, some sites do not have exceedances.

## Check variogram inside the simulation

```{r, message=FALSE}
x <- spa
y <- spa
t <- temp
RandomFields::RFoptions(spConform = FALSE, install = "no")
lx <- length(sx <- seq_along(x))  # spatial
ly <- length(sy <- seq_along(y))  # spatial
lt <- length(st <- seq_along(t))  # temporal
s0 <- c(1, 1)
t0 <- 1

## Model-Variogram BuhlCklu
modelBuhlCklu <- RandomFields::RMfbm(alpha = alpha1, var = 2*beta1, proj = 1) +
                  RandomFields::RMfbm(alpha = alpha1, var = 2*beta1, proj = 2) +
                  RandomFields::RMfbm(alpha = alpha2, var = 2*beta2, proj = 3)

## Construct grid
Nxy <- lx * ly # spatial grid size
N <- Nxy * lt # spatio-temporal grid size
grid <- matrix(0, nrow=N, ncol=3) # (N,3)-matrix

for (i in sx)
  for (j in seq_len(ly*lt))
    grid[i+(j-1)*ly, 1] <- i

for (i in sy)
  for (j in sx)
    for(k in st)
      grid[j+lx*(i-1)+(k-1)*Nxy, 2] <- i

for (i in st)
  for (j in seq_len(Nxy))
    grid[j+Nxy*(i-1), 3] <- i

# Construct shifted variogram for conditional spatio-temporal point
gamma <-  conditional_variogram(x, y, t, s0, t0, grid, modelBuhlCklu, adv)
```

```{r, fig.height=3, fig.width=3, fig.cap="Variogram relative to the conditional spatio-temporal point for all sites"}
par(mfrow = c(1, 1),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)
# Plot the variogram
plot(gamma, main = "", xlab = "Time lag",
    ylab = "Variogram")
```


```{r, fig.height=5, fig.width=5, fig.cap="Variogram relative to the conditional spatio-temporal point for four sites"}
par(mfrow = c(2, 2),
    cex = 0.7,
    pin = c(2, 1),
    lwd = 0.5,
    col = btfgreen)

# Plot the variogram for one site
plot(gamma[1,1, ], main = TeX(r"($\gamma(s-s_0, \tau)$ with $s=s_0=(1,1)$)"), xlab = "Time lag",
    ylab = "Variogram")
plot(gamma[2,1, ], main = TeX(r"($s=(2,1)$)"), xlab = "Time lag",
    ylab = "Variogram")
plot(gamma[5,2, ], main = TeX(r"($s=(5,2)$)"), xlab = "Time lag",
    ylab = "Variogram")
plot(gamma[5,5, ], main = TeX(r"($s=(5,5)$)"), xlab = "Time lag",
    ylab = "Variogram")
```


```{r, echo=TRUE, message=FALSE}
s0 <- c(1, 1)
s0_x <- s0[1]
s0_y <- s0[2]
t0 <- 1
print(gamma[s0_x, s0_y, t0] == 0)

s_x <- 1
s_y <- 1
time <- 2
semivario_s_t <- beta1 * abs(s0_x - s_x)^alpha1 + # no advection
                 beta1 * abs(s0_y - s_y)^alpha1 +
                 beta2 * abs(t0 - time)^alpha2
print(gamma[s_x, s_y, time] != 2*semivario_s_t) # as to be different

tau <- t0 - time
semivario_s_t_adv <- beta1 * abs(s0_x - s_x - adv[1] * tau)^alpha1 +
                 beta1 * abs(s0_y - s_y - adv[2]*tau)^alpha1 +
                 beta2 * abs(tau)^alpha2
print(gamma[s_x, s_y, time] == 2 * semivario_s_t_adv)

semivario_s_t_adv <- beta1 * abs(s0_x - s_x - adv[1] * abs(tau))^alpha1 +
                 beta1 * abs(s0_y - s_y - adv[2]*abs(tau))^alpha1 +
                 beta2 * abs(tau)^alpha2
print(gamma[s_x, s_y, time] == 2 * semivario_s_t_adv)


s_x <- 2
s_y <- 1
time <- 2
tau <- t0 - time
semivario_s_t_adv <- beta1 * abs(s0_x - s_x - adv[1] * tau)^alpha1 +
                 beta1 * abs(s0_y - s_y - adv[2]*tau)^alpha1 +
                 beta2 * abs(tau)^alpha2
print(gamma[s_x, s_y, time] == 2 * semivario_s_t_adv)

semivario_s_t_adv <- beta1 * abs(s0_x - s_x - adv[1] * abs(tau))^alpha1 +
                 beta1 * abs(s0_y - s_y - adv[2]*abs(tau))^alpha1 +
                 beta2 * abs(tau)^alpha2
print(gamma[s_x, s_y, time] == 2 * semivario_s_t_adv)


s_x <- 2
s_y <- 1
time <- 10
tau <- time - t0
semivario_s_t_adv <- beta1 * abs(s_x  - s0_x - adv[1] * tau)^alpha1 +
                 beta1 * abs(s_y - s0_y - adv[2]*tau)^alpha1 +
                 beta2 * abs(tau)^alpha2
print(gamma[s_x, s_y, time] == 2 * semivario_s_t_adv)

semivario_s_t_adv <- beta1 * abs(s_x - s0_x - adv[1] * abs(tau))^alpha1 +
                 beta1 * abs(s_y - s0_y - adv[2] * abs(tau))^alpha1 +
                 beta2 * abs(tau)^alpha2
print(gamma[s_x, s_y, time] == 2 * semivario_s_t_adv)
```

Problem: J'ai consideré seulement des tau positifs tels que tau=t-t0 avec $t\geq t0$.
Or il faut faire correspondre dans le bon sens les s-s0 et t-t0. Il faut donc que
mette s-s0 et non s0-s quand je calcule les lags spatio-temporels. FAIT.

Donc dans le cas des simulations de Brown-Resnick, sans site conditionel, je dois avoir le meme style de probleme avec 
des taus positifs ne correspondant pas à comment je prends les lags spatiales. Comment faire ?
On a $\tau = t_1 - t_2$ et $h=s_1 - s_2$ mais $\tau$ n'est pas toujours positifs... alors que je considère des $\tau \in \{1, 2, \ldots, 10\}$.
Idées: valeurs absolues des lags temporels dans le vaiogramme de la simu?
Considerer des tau négatifs dans l'optim dans le calcul des lags spatio-temporels?


